---
title: "Lung emergency presentation in Cosmos"
author: "Andrew Zimolzak"
date: "`r Sys.Date()`"
output:
  word_document: default
  pdf_document: default
---

```{r libs, include=FALSE}
library(dplyr)
library(knitr)
```

# Source data

```{r source-data}
emergency  <- c( 2189, 2714, 3306, 4082 )
patients <- c( 12445, 14680, 16429, 18579 )

race <- data.frame(
  factor_levels = c(
    'black', 'hawaii', 'indian-alaska', 'white', 'other', 'asian', 'none', 'hispanic'
  ),
  numerator = c(1796, 33, 76, 10177, 418, 161, 87, NA),
  denominator = c(7371, 139, 341, 52371, 2305, 1134, 1320, NA)
)

race_dropna <- race %>%
  filter(!is.na(numerator) & !is.na(denominator))

sex <- data.frame(
  factor_levels = c('male', 'female'),
  numerator = c(6148, 6221),
  denominator = c(30525, 32542)
)

age <- data.frame(
  factor_levels = c('<69', '69-77.9', '78+'),
  numerator = c(4991, 3818, 3561),
  denominator = c(21753, 20537, 20779)
)

per_cap_income_zip <- data.frame(
  factor_levels = c('<23', '23-26.9', '27-30.9', '31-39.9', '40+'),
  numerator = c(2508, 2558, 2688, 2930, 1570),
  denominator = c(11380, 12009, 12625, 15491, 10664)
)

```


# Hypotheses tests by factor

```{r function-def, include=FALSE}
rows2fisher <- function(positives, totals) {
  negatives = totals - positives
  rows = length(positives)
  M = matrix(c(positives, negatives), nrow = rows)
  return(fisher.test(M, simulate.p.value = TRUE))
}
```


## Vulnerabilty index

This is the only one where I chose to do trend test.

```{r vuln}
prop.trend.test(emergency, patients)
rows2fisher(emergency, patients)
```

## Race
```{r}
prop.test(race$numerator, race$denominator)
rows2fisher(race_dropna$numerator, race_dropna$denominator)
```

## Sex
```{r}
prop.test(sex$numerator, sex$denominator)
rows2fisher(sex$numerator, sex$denominator)
```

## Age
```{r}
prop.test(age$numerator, age$denominator)
rows2fisher(age$numerator, age$denominator)
```

## Income

Note it is specifically *estimated per-capita income,*  and this is on the ZIP code level. Meaning everyone in the same ZIP will have the same income assigned.

```{r}
prop.test(per_cap_income_zip$numerator, per_cap_income_zip$denominator)
rows2fisher(per_cap_income_zip$numerator, per_cap_income_zip$denominator)
```




# Re-create table 1

Table 1 in the VA paper goes as follows: white, black, (hispanic), asian, pacific, ai/an, other+unknown.

```{r table1}
paper_order <- c(
  'white', 'black', 'hispanic', 'asian', 'hawaii', 'indian-alaska', 'other', 'none'
)
row_nums_in_order <- match(paper_order, race$factor_levels)

race[row_nums_in_order, ] %>%
  mutate(percent = round(100 * numerator / denominator, 2)) %>%
  kable()

sex %>% mutate(percent = round(100 * numerator / denominator, 2)) %>% kable()

age %>% mutate(percent = round(100 * numerator / denominator, 2)) %>% kable()
```




# Sanity checks

About total counts, which might not equal each other, but which should be close:

```{r sanity}
sum(race$denominator, na.rm = TRUE)
sum(age$denominator)
sum(sex$denominator)
```

How many to aim for in a quintile of income:

```{r sanity2}
round(sum(age$denominator) / 5)
```
